---
title: "loopR: Gallery of examples"
author: "Michael Kammer"
date: "`r Sys.Date()`"
output: 
    rmarkdown::html_vignette:
        fig_width: 8
        fig_height: 6
        toc: true
        toc_depth: 2
        number_sections: true
vignette: >
  %\VignetteIndexEntry{loopR: Gallery of examples}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
bibliography: "Demo_references.bib"
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(magrittr)
library(purrr)
library(loopR)
```

This vignette contains example annotated use-cases of the `loopR` package.
We introduce several of the available options and briefly explain their effect.
For an in-depth introduction into the goals of this package and the general 
workflow, see the vignette ["loopR: A package for creating nested loop plots"](Demo.html).

For many options to further customize the plots, please refer to the 
documentation of the underlying [ggplot2 package](https://ggplot2.tidyverse.org/index.html).

# Example 1: general use-case
This example uses the same artifically created dataset as the other vignette. 

## Data
```{r echo=TRUE}
set.seed(14)
params = list(
    samplesize = c(100, 200, 500),
    param1 = c(1, 2), 
    param2 = c(1, 2, 3), 
    param3 = c(1, 2, 3, 4)
)

design = expand.grid(params)

# add some "results"
design %<>% 
    mutate(method1 = rnorm(n = n(),
                           mean = param1 * (param2 * param3 + 1000 / samplesize), 
                           sd = 2), 
           method2 = rnorm(n = n(),
                           mean = param1 * (param2 + param3 + 2000 / samplesize), 
                           sd = 2), 
           method3 = rnorm(n = n(),
                           mean = param1 * (param2 + param3 + 3000 / samplesize), 
                           sd = 2))

print(head(design, n = 10))
```

## Basic facetted nested loop plot {#basic}

- Define x-axis, grid facets and parameter step functions (`x`, `grid_rows`, 
`grid_cols`, `steps`). 
- Adjust step positions (`steps_y_base`, `steps_y_height`).
- Set axis names (`x_name`, `y_name`).
- Adjust spacing between connected areas on the x-axis (`spu_x_shift`).
- Set color specifications via a function (`colors`).
- Adjust step annotations (`steps_values_annotate`, `steps_annotation_size`).
- Add a horizontal line with intercept 0 (`hline_intercept`).
- Expand the y-axis for proper display of the parameter step functions 
(`y_expand_add`).
- Rotate x-axis labels and adjust their position and size (`add_custom_theme`).

```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## Only rows

- Only set `grid_rows` argument, move other parameters into `steps`.
- Re-ajust steps positions.
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = c("param2", "param3"),
                     grid_rows = "param1", 
                     steps_y_base = -10, steps_y_height = 3, steps_y_shift = 3,
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## Only columns
- Only set `grid_cols` argument, move other parameters into `steps`.
- Re-ajust steps positions.
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = c("param2", "param3"),
                     grid_cols = "param1", 
                     steps_y_base = -5, steps_y_height = 1, steps_y_shift = 3,
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## "Classic" nested loop plots

- Move all parameters except one into `steps`.
- Connect all the results via step functions: set argument `draw` to 
`c("add_points", "add_steps")` to display results via points and steps, then 
set `connect_spus` to `TRUE`.
- Re-ajust steps positions.

```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = c("param1", "param2", "param3"),
                     draw = "add_steps", 
                     steps_y_base = -5, steps_y_height = 1, steps_y_shift = 3,
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 200,
                     connect_spus = TRUE,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 5) 
                         )
                     ))
print(p)
```

## Spacing on x-axis
So far the x-axis has been treated as numeric scale with spacing of the breaks
according to the values of the design parameter which defines the x-axis. 
If the design parameter is actually a factor, or converted to be one, 
then the spacing on the x-axis is equal.

- Input data.frame for `nested_loop_plot` now uses a factor as x-axis 
(converted via `dplyr::mutate`).
- Adapt gap between different connected areas of the plot (spu) via 
`spu_x_shift`. Note that the gap between any two datapoints on the x-axis is 
now exactly one unit.
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
x_factor_design = design %>% 
    mutate(samplesize = as.factor(samplesize))
p = nested_loop_plot(resdf = x_factor_design, 
                     x = "samplesize", steps = c("param1", "param2", "param3"),
                     draw = "add_steps", 
                     steps_y_base = -5, steps_y_height = 1, steps_y_shift = 3,
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 1,
                     connect_spus = TRUE,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 5) 
                         )
                     ))
print(p)
```

## Re-labelling data
Data can be re-labelled to streamline the display of parameters on the x-axis
or give meaningful names to panels.

- Relabel data using `replace_labels`.
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     replace_labels = list(
                         samplesize = c("100$" = "100", 
                                        "200$" = "", 
                                        "500$" = "500"), 
                         param2 = c("1" = "low", 
                                    "2" = "mid", 
                                    "3" = "high"),
                         param3 = c("1" = "very low",  # partial replacement
                                    "4" = "very high")
                     ),
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## Changing and adding design elements
Elements of the plot can be customized. Note that these parameters 
(`sizes`, `line_linetypes`, `point_shapes`, `colors`) all at least take either
a single numeric value, a vector of numeric values or a named vector of 
numeric values specifying which method should have which value.

- Provide different linetpyes per method (`line_linetypes`).
- Provide different point shapes per method (`point_shapes`).
- Provide different point sizes per method. This requires passing a numeric
vector to `sizes` and setting `point_size` to `NULL` due to the underlying 
implementation in the `ggplot2` package.
- Fixing linewidth to 1 (`line_size`).
- Add a ribbon around zero, .e.g to indicate confidence regions, simulation
monte carlo error or a target region the methods should reach (added to 
post-processing list via `add_annotation`).
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     line_linetypes = c(1, 2, 3),
                     point_shapes = c("method1" = 2, "method2" = 3, "method3" = 1),
                     sizes = c(1, 1.5, 2), point_size = NULL, 
                     line_size = 1,
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         # add ribbon via an annotation
                         add_annotation = list(
                             geom = "rect",
                             xmin = -Inf, xmax = Inf, # stretch whole x-axis
                             ymin = 0, ymax = 10, alpha = 0.1, fill = "black"
                         ),
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```


## Partial designs
In a partial design, not every possible combination of values for the 
design parameters is present in the results from the experiment. In our 
example, we simply remove parts corresponding to two design parameters 
from the data.frame containing the data. 

- The `design_type` can be set to "partial" to remove the parameter step 
functions for the missing design parts. If left at default ("full"), then 
the missing parts are added. 

```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
partial_design = design %>% 
    filter(!(param2 == 2 & param3 %in% c(3, 4)))
p = nested_loop_plot(resdf = partial_design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     design_type = "partial",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

An alternative parameter to use in case of missing data due to partial designs
determines how "holes" in the data are handled. 

- The `na_rm` argument can be set to TRUE (default) to ignore missing data
when connecting result values. (Note that in the middle panel no data is 
available for samplesize 200.)
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
partial_design = design %>% 
    filter(!(param2 == 2 & samplesize == 200))
p = nested_loop_plot(resdf = partial_design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     na_rm = TRUE, 
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

- The `na_rm` argument can be set to FALSE make it clear that data is missing 
by not connecting the result values accross missing data.
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
partial_design = design %>% 
    filter(!(param2 == 2 & samplesize == 200))
p = nested_loop_plot(resdf = partial_design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     na_rm = FALSE, 
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## Free axes and axis adjustments
Axes of the facets can have different scales. However, currently it is only
possible to have different y- and x-scales for each row and column, respectively. 
This is made possible thanks to the [`facet_grid_sc`](https://github.com/zeehio/facetscales) 
extension of `ggplot2`. Because of this, it is also possible to adjust the 
axes individually (i.e. per row / per column). 

- Set `grid_scales` argument to "free_y" to free the y-scale. The x-axis remains
the same, even if set to be free.
- Adjust y-axis only in second row by passing a list to `y_expand_add` (entries
of this list are named by the values of the design parameter which defines
the `grid_rows`).
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     grid_scales = "free_y",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = list(
                         "2" = c(10, NULL)
                     ),
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## Addtional data display
The package supports adding steps for meta-information. These do not influence
the layout of the plot. 

- Additional steps are added via `steps_add` (note that additional column 
added to the design data.frame).
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
meta_design = design %>% 
    mutate(Difficulty = if_else(param3 == 4, "Hard", "Easy"))
p = nested_loop_plot(resdf = meta_design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_add = "Difficulty",
                     steps_y_base = -10, steps_y_height = 5, steps_y_shift = 10, 
                     x_name = "Samplesize", y_name = "Error",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## Adding geometries to the plot
Further geometries can be added to the plot via post-processing. Here we add
text to annotate the results, but also additional points or lines could be 
drawn in the same way. By passing a new data.frame, even new data can be 
added in this step, but this is best done using the modular approach 
[shown below](#modular).

- Annotate results by adding labels to them in post-processing (`add_text`).
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 5, 
                     x_name = "Samplesize", y_name = "Error",
                     draw = "add_lines",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     post_processing = list(
                         add_text = list(
                             decimals = 1, size = 2, vjust = 0, hjust = 0
                         ),
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

## Axis transformations

The normal axis transformations of `ggplot2` can not be used in general as the
steps added for simulation parameters are also on the same y-scale as the 
simulation results. E.g. if a logarithmic scale is to be used for results, but 
the parameter steps have negative y-coordinates, then the approach using `ggplot2`
will not work.

Hence, the functionality is implemented by transformation of the data. 
This can be done by the user outside the `loopR` package or passed to the
functions implemented in this package using the `trans` argument.

Note that this is a transformation of the data. Hence, the y-axis labels are
also on the transformed scale. If that is not wanted, the user can easily 
reverse this transformation for the labels by passing a function to the 
`y_labels` argument. See documentation. Another possibility is to simply
rename the y-axis as e.g. 'log2 of results' to indicate the scaling. 

A further consequence is that all additions to the plot such as lines or 
parameter steps are now also on the transformed scale and may require
re-adjustment.

- Define transformation by setting `trans`. 
- Set `ylim`, `y_breaks` and `y_labels` arguments.
- Adjust parameter step functions and y-axis expansion for proper display.
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = "param3",
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -0.5, steps_y_height = 0.5, 
                     x_name = "Samplesize", y_name = "Error",
                     trans = log2,
                     ylim = c(0, 6), 
                     y_breaks = 0:6,
                     y_labels = function(x) ifelse(x > 0, 2^x, 0),
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(3, 0.5), 
                     post_processing = list(
                         add_custom_theme = list(
                             axis.text.x = element_text(angle = -90, 
                                                        vjust = 0.5, 
                                                        size = 8) 
                         )
                     ))
print(p)
```

# Example 2: many design parameter

## Data
We expand the design parameters by another layer of parameters. Overall the 
setup contains 600 unique parameter combinations.

```{r}
set.seed(82356)
params = list(
    samplesize = c(10, 50, 100, 200, 500),
    param1 = c(1, 2), 
    param2 = c(1, 2, 3), 
    param3 = c(1, 2, 3, 4), 
    param4 = c(1, 2, 3, 4, 5)
)

design = expand.grid(params)

design %<>% 
    mutate(method1 = rnorm(n = n(),
                           mean = param1 * (param2 * param3 * param4 + 50 / samplesize), 
                           sd = 2), 
           method2 = rnorm(n = n(),
                           mean = param1 * (param2 + param3 + param4 + 100 / samplesize), 
                           sd = 2), 
           method3 = rnorm(n = n(),
                           mean = param1 * (param2 + param3 * param4 + 150 / samplesize), 
                           sd = 2))
```

## Plots

- Set x-axis labels to NULL to avoid cluttering display (`x_labels`). Add
values on x-axis to name of axis.
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = c("param3", "param4"),
                     grid_rows = "param1", grid_cols = "param2", 
                     steps_y_base = -10, steps_y_height = 2.5, steps_y_shift = 5, 
                     x_name = "Samplesize (100, 200, 500)", y_name = "Error",
                     draw = "add_lines",
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     x_labels = NULL,
                     post_processing = list(
                         add_custom_theme = list(legend.position = "top")
                     ))
print(p)
```

- Add grid lines as in @NestedLoop (adjust theme in `add_custom_theme`).
```{r message=FALSE, warning=FALSE, echo=TRUE, results="hide"}
p = nested_loop_plot(resdf = design, 
                     x = "samplesize", steps = c("param1", "param2", "param3", "param4"),
                     steps_y_base = -10, steps_y_height = 2.5, steps_y_shift = 7.5, 
                     x_name = "Samplesize (100, 200, 500)", y_name = "Error",
                     draw = "add_lines", 
                     spu_x_shift = 75,
                     colors = scales::brewer_pal(palette = "Dark2"),
                     steps_values_annotate = TRUE, steps_annotation_size = 2.5, 
                     hline_intercept = 0, 
                     y_expand_add = c(10, NULL), 
                     x_labels = NULL,
                     post_processing = list(
                         add_custom_theme = list(
                             legend.position = "top", 
                             panel.grid.major = element_line(
                                 color = "grey95", size = 0.1
                             )
                         )
                     ))
print(p)
```


# Modular use of the package {#modular}
While the main interface to this package is a single function, the creation
of nested loop plots is actually implemented via modular functions. 
This is useful if the user wants more control over the design of the plot
and especially if additional meta-data should be added. 

## Step-by-step
We recreate the [basic facetted nested loop plot](#basic) using the 
same data step by step. 
These are

- Creation of basic plotting data via `nested_loop_base_data`. This data.frame
contains all information to create the nested loop plots from this package, 
except for the data for drawing design parameter step functions. It 
can be used to add further information for plotting.
- Creation of a basic `ggplot2` object via `nested_loop_base_plot`. It does not
contain design parameter step functions.
- Addition of data for drawing parameter step functions via `nested_loop_paramsteps_data`.
- Updating the `ggplot2` object for drawing parameter step functions.
- Add post-processing to the `ggplot2` object to adjust scales and axes.

```{r include=FALSE}
set.seed(14)
params = list(
    samplesize = c(100, 200, 500),
    param1 = c(1, 2), 
    param2 = c(1, 2, 3), 
    param3 = c(1, 2, 3, 4)
)

design = expand.grid(params)

# add some "results"
design %<>% 
    mutate(method1 = rnorm(n = n(),
                           mean = param1 * (param2 * param3 + 1000 / samplesize), 
                           sd = 2), 
           method2 = rnorm(n = n(),
                           mean = param1 * (param2 + param3 + 2000 / samplesize), 
                           sd = 2), 
           method3 = rnorm(n = n(),
                           mean = param1 * (param2 + param3 + 3000 / samplesize), 
                           sd = 2))
```

```{r message=FALSE, warning=FALSE}
plot_data = nested_loop_base_data(
    design, 
    x = "samplesize", steps = "param3",
    grid_rows = "param1", grid_cols = "param2",
    spu_x_shift = 75
)

p = nested_loop_base_plot(
    plot_data,
    x_name = "Samplesize",
    y_name = "Error", 
    colors = scales::brewer_pal(palette = "Dark2")
)
print(p)

plot_data = nested_loop_paramsteps_data(
    plot_data,
    steps_y_base = -10,
    steps_y_height = 5
)

p = nested_loop_paramsteps_plot(
    p, plot_data, 
    steps_values_annotate = TRUE, 
    steps_annotation_size = 2.5
)
print(p)

p = add_processing(
    p, 
    list(
        # set limits
        adjust_ylim = list(
            y_expand_add = c(10, NULL)
        ),
        # adjust theme
        add_custom_theme = list(
            axis.text.x = element_text(angle = -90, 
                                       vjust = 0.5, 
                                       size = 8)
        ), 
        # add horizontal lines
        add_abline = list(
            intercept = 0
        )
    )
)
print(p)
```

## Advanced usage
Using the package functions modularly opens up many more options for 
customization. These are all based on modification of the data.frames which 
are returned by `nested_loop_base_data` and `nested_loop_paramsteps_data` that
provide the basis data for the plotting functions of this package.

We demonstrate an exemplary customization here. First, we obtain the 
necessary data. The data.frame of interest is stored in the `plotdf` entry of
the resulting list, which can be modified to add meta-information to the 
plots. We do this using the `dplyr::mutate` and `dplyr::case_when` 
functionality.

- Annotate scenarios of special interest, e.g. due especially interesting
combination of design parameters. Here we indicate scenarios with samplesize
of 100 or 200, param3 of 1 and param2 either 2 or 3.
- Annotate unusually high values by their numeric value. 

```{r message=FALSE, warning=FALSE}
plot_data = nested_loop_base_data(
    design, 
    x = "samplesize", steps = "param3",
    grid_rows = "param1", grid_cols = "param2",
    spu_x_shift = 75
) %>%
    nested_loop_paramsteps_data(
        plot_data,
        steps_y_base = -10,
        steps_y_height = 5
    )

# create annotation data
plot_data$plotdf %<>% 
    mutate(
        annotate_scenario = case_when(
            samplesize %in% c(100, 200) & 
                param3 == 1 & 
                param2 %in% c(2, 3)
                ~ 0 # height at which a point should be drawn on y-axis
            ),
        annotate_value = case_when(y_coord > 60 ~ y_coord) %>% 
            round(2)
    )
```

We add the additional plot elements to the post-processing list via `add_points`
and `add_text`.
```{r message=FALSE, warning=FALSE}
p = nested_loop_base_plot(
    plot_data,
    x_name = "Samplesize",
    y_name = "Error", 
    colors = scales::brewer_pal(palette = "Dark2")
) %>% 
    nested_loop_paramsteps_plot(
        plot_data, 
        steps_values_annotate = TRUE, 
        steps_annotation_size = 2.5
    ) %>% 
    add_processing(
        list(
            # annotate interesting scenarios
            add_points = list(
                mapping = aes(y = annotate_scenario), 
                shape = 8, color = "red"
            ),
            # annotate high values
            add_text = list(
                mapping = aes(label = annotate_value), 
                color = "black", size = 3, hjust = 0, vjust = 0,
                nudge_x = 20, nudge_y = 1
            ),
            # set limits
            adjust_ylim = list(
                y_expand_add = c(10, NULL)
            ),
            # adjust theme
            add_custom_theme = list(
                axis.text.x = element_text(angle = -90, 
                                           vjust = 0.5, 
                                           size = 8)
            )
        )
    )
print(p)
```

# R session information {#rsession}
```{r echo=FALSE}
sessionInfo()
```

# References